---
title: "EDA strawberry"
author: "Jiun Lee"
date: "2022-11-07"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(readxl)
strawb <- read_xlsx("strawberries-2022oct30-a.xlsx",col_names = TRUE)
```
##This document got a reference from ag_data_introduction.R by Haviland. I followed the instruction in the file and added my codes for cleaning the strawberry dataset.

```{r}
## Get the column names and index them
cnames <- colnames(strawb) 
x <- 1:dim(strawb)[2]
T <- NULL

## Collect number of unique rows in each column
for(i in x){T <- c(T, dim(unique(strawb[i]))[1])}
drop_cols <- cnames[which(T == 1)]
strawb %<>% select(!all_of(drop_cols))
## Let's arrange the data frame by year and state.
strawb %<>% arrange(Year, State)

##clean the data.
strawb %<>% separate(col=`Data Item`,
                    into = c("Strawberries", "type", "items", "units"),
                    sep = ",",
                    fill = "right")
```

## let's track down the organic entries
```{r}
type_organic <- grep("organic", 
               strawb$type, 
               ignore.case = T)

items_organic <- grep("organic", 
                      strawb$items, 
                      ignore.case = T)  ## nothing here

Domain_organic <- grep("organic", 
                      strawb$Domain, 
                      ignore.case = T)

Domain_Category_organic <- grep("organic", 
                       strawb$`Domain Category`, 
                       ignore.case = T)

## OK.  So let's create a strawb_organic tibble
## All three are the same

same <- (intersect(type_organic, Domain_organic)==
         intersect(type_organic, Domain_Category_organic))
length(same)==length(type_organic)

org_rows <- intersect(type_organic, Domain_organic)

strawb_organic <- strawb %>% slice(org_rows, preserve = FALSE)

strawb_non_organic <- strawb %>% filter(!row_number() %in% org_rows)
```

## now separate the chemical data
```{r}
temp1 <- strawb_non_organic %>% select(type) %>% 
  distinct()

## the chemical data in the column "type" 
## is in the rows where type = BEARING - Applications

chem_rows <- grep("BEARING - APPLICATIONS", 
                       strawb_non_organic$type, 
                       ignore.case = T)
## of the 2172 rows in strawb_non_organic 2112 contain
## chemical data

## the column "Domain" also contains indications of rows 
## 

chem_rows_1 <- grep("chemical", 
                  strawb_non_organic$Domain, 
                  ignore.case = T)

## there are less chem rows in the Domain column
## than in the type column

ins <- intersect(chem_rows, chem_rows_1)

## the cardinality of the intersection equals the cardinality
## of chem_rows_1.  So, chem_rows calls all of the chem rows 
## we have found so far.

## Finally, we examine the `Domain Category`column.

chem_rows_2 <- grep("chemical", 
                    strawb_non_organic$`Domain Category`, 
                    ignore.case = T)

ins_2 <- intersect(chem_rows, chem_rows_2)

## again, all of the row numbers in chem_rows_2 are in 
## chem_rows


## now sort out these tibbles just like we did strawb
## Now, create a chem tibble


strawb_chem <- strawb_non_organic %>% slice(chem_rows, preserve = FALSE)

## now clean up the workspace before tackling the three tibbles just created.

rm(x, T, drop_cols, temp1, r_thiram, r_thiram_1,
   df_carbendazim, df_Bifenthrin, df_methyl_bromide, 
   df_1_3_dichloropropene, df_chloropicrin, df_Telone,
   pr_rec, type_organic, items_organic, Domain_organic,
   Domain_Category_organic, same, org_rows, chem_rows,
   chem_rows_1, chem_rows_2, ins, ins_2, cnames, i)
```
## now the environment lists the tibbles strawb, strawb_chem,
## strawb_non_organic, and strawb_organic

## first drop the "no-info" columns  Since we will need to do this 
## for all three tibbles, write a functin
```{r}

  before_cols = colnames(strawb_chem)
  T = NULL
  x = length(before_cols)

   for(i in 1:x){
    b <- length(unlist(strawb_chem[,i] %>% unique()))
    T <- c(T,b)
   }
    
  drop_cols <- before_cols[which(T == 1)]
  strawb_chem %<>% select(!all_of(drop_cols))
  after_cols = colnames(strawb_chem)

## drop_cols is labeling information
  
  temp1 <- strawb_chem %>% select(units) %>% distinct()
  ## in units rows are either NA or AVG
  
## separate Domain Category at :
  
  strawb_chem %<>% separate(col=`Domain Category`, 
                                    into = c("dc1", "chem_name"),
                                    sep = ":", 
                                    fill = "right")
  
temp1 <- strawb_chem %>% select(chem_name) %>% unique()
length(unlist(temp1))

## first clean it up

aa  <- grep("measured in", 
                    strawb_chem$items, 
                    ignore.case = T)
length(aa)

## so "item" is really units

## are Domain and dc1 the same thing?

sum(strawb_chem$Domain == strawb_chem$dc1) == dim(strawb_chem)[1]

## yes, they are. let's work with dc1. Also, drop State ANSI

strawb_chem %<>% select(Year, State, items, units, dc1, chem_name, Value)

## now you can see that this tibble needs to be wider.  
## clean up the entries before using pivot_wider

## items should only contain unit names
## rename units to category

strawb_chem %<>% rename(category = units)

## remove "MEASURED IN " 

strawb_chem$items <- str_remove_all(strawb_chem$items, "MEASURED IN ")

## rename items to units
strawb_chem %<>% rename(units = items)

## Do all the dc1 entries begin with "Chemical"?

bb  <- grep("CHEMICAL, ", 
            strawb_chem$dc1, 
            ignore.case = T)
length(bb)
chem <- 1:2112

non_chem_rows <- setdiff(chem, bb)
length(non_chem_rows)

## on let's look at these rows in a tibble

temp1 <- strawb_chem %>% slice(non_chem_rows)

### !! fertilizers  

## keep them -- probably won't use them as a lone tibble

fertilizers <- temp1

## cleanup
rm(temp1, temps, temp3, aa, bb)

## now remove "CHEMICAL, " from the entries in the dc1
## and rename the column chem_types


strawb_chem$dc1 <- str_remove_all(strawb_chem$dc1, "CHEMICAL, ")

strawb_chem$dc1 %>% unique()

strawb_chem %<>% rename(chem_types = dc1)

## now fix the chem_name column
## remove the parens


strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\(")
strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\)")


## separate chem_name and chem_code

strawb_chem %<>% separate(col = chem_name,
                          into = c("chem_name","chem_code"),
                          sep = "=",
                          fill = "right") 
```

## now fill in a label fot NA in the category column
## first check that "lb" in the units column corresponds 
## to NA in the category column
```{r}
aa <- which(strawb_chem$units == " LB")

bb <- which(is.na(strawb_chem$category))

sum(aa==bb)==length(aa)

## yes, they correspond
## So, we can say all units except 'LB' unit corresponds to 'AVG'

## The category column is meaningless. Let's delete the category column.
strawb_chem <- strawb_chem %>% select(-category)


##The chem_code column seems meaningless.
strawb_chem <- strawb_chem %>% select(-chem_code)
##Let's see the if there's NA in chem_code
##First, remove () in colum Value.
strawb_chem$Value <- str_remove_all(strawb_chem$Value, "\\(")
strawb_chem$Value <- str_remove_all(strawb_chem$Value, "\\)")

##Let's remove NA in column Value.
aa <- which(strawb_chem$Value == "NA")
length(aa)
chem <- 1:2112
cc <- setdiff(chem, aa)
length(cc)
strawb_chem <- strawb_chem %>% slice(cc)
#Now, the obs of strawb_chem is 1878.

##Let's remove D in column Value.
aa <- which(strawb_chem$Value=="D")
chem <- 1:1878
cc <- setdiff(chem, aa)
strawb_chem <- strawb_chem %>% slice(cc)

##Now, the strawb_chem dataset's observation is 900.

##What's the type of column Value?
typeof(strawb_chem$Value)

##It's character.
##Let's change it to numeric.
strawb_chem$Value <- as.numeric(strawb_chem$Value)

##NAs introduced. Let's remove it.
aa <- which(is.na(strawb_chem$Value))
chem <- 1:900
cc <- setdiff(chem, aa)
strawb_chem <- strawb_chem %>% slice(cc)

##clean the data.
rm(temp1,aa, after_cols,before_cols,cc,chem,drop_cols,T,x,b,i)

##Let's go back to column units.
unique(strawb_chem[3]) 
##values are "LB", "LB / ACRE / APPLICATION","LB / ACRE / YEAR".

##Seperate the datasets by units.

##grep the row numbers of "LB / ACRE / APPLICATION"
aa <- grep("APPLICATION",strawb_chem$units,ignore.case = T)

##dataset of "LB / ACRE / APPLICATION" unit
chem_app <- strawb_chem %>% slice(aa, preserve=F)
chem_app <- chem_app %>% select(-units)

##grep the row numbers of "LB / ACRE / YEAR"
aa <- grep("YEAR",strawb_chem$units,ignore.case = T)

##dataset of "LB / ACRE / YEAR" unit
chem_year <- strawb_chem %>% slice(aa, preserve=F)
chem_year <- chem_year %>% select(-units)


##grep the row numbers of "LB"
aa <- grep("^ LB$",strawb_chem$units,ignore.case = T)

##dataset of "LB" unit
chem_lb <- strawb_chem %>% slice(aa, preserve=F)
chem_lb <- chem_lb %>% select(-units)
```

##There are some safe chemicals. Let's get the dataset of them.
```{r cars}
#find a row numbers of notorious chemicals
bb  <- grep("BIFENTHRIN", 
            strawb_chem$chem_name, 
            ignore.case = T)
b <- strawb_chem %>% slice(bb, preserve=F)
#chemtype:insecticide

cc <- grep("Carbendazim", 
            strawb_chem$chem_name, 
            ignore.case = T)  ##empty

mm <- grep("MENTHYL BROMIDE", 
            strawb_chem$chem_name, 
            ignore.case = T)  ##empty

dd <- grep("DICHLOROPROPENE", 
            strawb_chem$chem_name, 
            ignore.case = T)
d <- strawb_chem %>% slice(dd, preserve=F)
#chemtype:other

tt <- grep("Telone", 
            strawb_chem$chem_name, 
            ignore.case = T) ##empty

ch <- grep("CHLOROPICRIN", 
            strawb_chem$chem_name, 
            ignore.case = T)
chc <- strawb_chem %>% slice(ch, preserve=F)
#chemtype : other

#combine them.
danger <- c(bb,ch,dd)
#get a dataset except the rows of dangerous chemicals
chem_x_danger <- strawb_chem %>% filter(!row_number() %in% danger)
rm(dd,ch,danger)
```

##Let's get the dataset of sales for organic
```{r pressure, echo=FALSE}
##rownumbers of sales type column = rownumbers of items not having sales (all rows - rows that having sales)

#units column = NA -> total of sales(CWT,$)
total <- which(is.na(strawb_organic$units))
sales_organic <- strawb_organic %>% filter(!row_number() %in% total)

cnames <- colnames(sales_organic) 
x <- 1:dim(sales_organic)[2]
T <- NULL

## Collect number of unique rows in each column
for(i in x){T <- c(T, dim(unique(sales_organic[i]))[1])}
## Use T to select columns to drop -- 
drop_cols <- cnames[which(T == 1)]
## Now, drop the columns with only one unique value.
sales_organic %<>% select(!all_of(drop_cols))
##Let's remove the column State ANSI
sales_organic <- sales_organic %>% select(-'State ANSI')

## remove "MEASURED IN " 
sales_organic$units <- str_remove_all(sales_organic$units, "MEASURED IN ")

##rename
sales_organic <- sales_organic %>% rename(units=items)

##seperate items
sales_organic %<>% separate(col=`items`, 
                                    into = c("items", "sales"),
                                    sep = "-", 
                                    fill = "right")
##remove sales column
sales_organic %<>% select(-sales)
## Let's arrange the data frame by units
sales_organic %<>% arrange(units)

#remove (D)
sales_organic <- sales_organic[-which(sales_organic$Value == "(D)"),]

```


##dataset of sales for unorganic?
```{r pressure, echo=FALSE}
#Let's get the non-chemical non-organic dataset.
chem_rows <- grep("BEARING - APPLICATIONS", 
                       strawb_non_organic$type, 
                       ignore.case = T)
nonorg_x_chem <- strawb_non_organic %>% filter(!row_number() %in% chem_rows)

##nonorg_x_chem doesn't look meaningful.
##It means, we can consider only 'strawb_chem' dataset as nonorganic dataset.

```

